Predicting Bike Usage at UofT St. George with Bayesian Poisson Regression

### 2024-12-01 paper/paper.qmd
More LLM claims，prose checked
More LLM claims，prose checked

## Overview
Bike-sharing services have experienced rapid growth in the Toronto region, with ridership rising dramatically from approximately 665,000 trips in 2015 to over 4.5 million in 2022, according to a study by the University of Toronto's School of Cities[@liu2023bike]. This rapid increase highlights the critical need for data-driven analyses to optimize bike-sharing systems and guide policy decisions that can further enhance urban mobility[@xiamen]. While city-wide studies provide valuable suggestions, research from Hangzhou, China, has demonstrated that bike-sharing usage patterns can vary significantly between urban areas and university campuses[@hangzhou]. These differences suggest that general urban studies conducted in the Toronto region may not fully capture the unique dynamics of campus-based bike-sharing systems. Therefore, to address this gap in the literature, this study focuses on the University of Toronto St. George Campus as a case study, analyzing bike-sharing usage within the campus in Toronto to provide a nuanced understanding of its role in campus mobility and inform policies tailored to similar settings.

@@

This data is particularly ideal for the study also due to its high reliability, which stems from the commercial nature of bike-sharing operations. The advanced bike-sharing system ensures accurate data collection and management by automatically recording ride information. Each bike and docking station is equipped with sensors and communication devices, enabling the capture of detailed trip data.

@@

start |>
  tail() |>
  kable(booktabs = TRUE, align = "c")
```

@@

@tbl-varibale is the summary of varibales. It is evident that the dataset includes 27 unique station names and spans nearly eight years (from January 1, 2017, to September 30, 2024), providing opportunities to analyze long-term trends and seasonal variations. In the count column, the minimum and median values are both 1, the mean is 2, the first quartile is 1, the third quartile is 3, and the maximum reaches 31, showing a significant right-skewed distribution. This indicates that most of the records have low count values, but there are a few extreme high values.
```{r}
#| warning: false
#| message: false
#| echo: false
#| label: tbl-varibale
#| tbl-cap: "Variables' summary"
count_min <- min(start$count)
count_1st_qu <- quantile(start$count, 0.25)
count_median <- median(start$count)
count_mean <- mean(start$count)
count_3rd_qu <- quantile(start$count, 0.75)
count_max <- max(start$count)

@@

start_asdate <- start 
start_asdate$time <- as.Date(start_asdate$time)
daily_counts <- start_asdate %>%
  group_by(time) %>%
  summarise(total_count = sum(count))

lm_model <- lm(total_count ~ time, data = daily_counts)

slope <- coef(lm_model)["time"]

@@

start_astime <- start 
start_astime$time <- as.POSIXct(start_astime$time)

start_astime <- start_astime %>%
  mutate(time_interval = case_when(
    hour(time) >= 0 & hour(time) < 4 ~ "00:00 - 04:00",
    hour(time) >= 4 & hour(time) < 8 ~ "04:00 - 08:00",
    hour(time) >= 8 & hour(time) < 12 ~ "08:00 - 12:00",
    hour(time) >= 12 & hour(time) < 16 ~ "12:00 - 16:00",
    hour(time) >= 16 & hour(time) < 20 ~ "16:00 - 20:00",
    hour(time) >= 20 & hour(time) < 24 ~ "20:00 - 00:00"
  ))

total_counts <- start_astime %>%
  group_by(time_interval) %>%
  summarise(count = sum(count, na.rm = TRUE)) %>%
  mutate(percentage = count / sum(count) * 100)

@@

average_counts_start <- start_1day1 %>%
  group_by(station_name) %>%
  summarize(average_count = mean(total_count, na.rm = TRUE))

average_counts_stop <- stop_1day1 %>%
  group_by(station_name) %>%
  summarize(average_count = mean(total_count, na.rm = TRUE))

@@

merged_data <- merged_data %>%
  mutate(count_difference = ifelse((average_count.y - average_count.x) >= 0, 
                                   ceiling(average_count.y - average_count.x), 
                                   floor(average_count.y - average_count.x)))

@@

This study considered using the ARIMA (autoregressive integrated moving average) model as an alternative but ultimately decided against it due to its limitations in capturing the specific characteristics of bike-sharing usage data. Although the ARIMA model has advantages in time series analysis, allowing for more accurate short-term predictions with parameters that have clearer interpretability. Bike-sharing usage exhibits significant seasonal patterns, time-of-day effects, and count-based characteristics, while ARIMA models are primarily suited for stationary time series with linear autocorrelations. ARIMA struggles to effectively address non-stationary trends and overdispersion, such as the marked decrease in usage during winter. Although extensions like SARIMA (Seasonal ARIMA) can account for seasonality, they still cannot directly handle the count-based nature of the data. In contrast, Bayesian Poisson regression is specifically designed for count data, allowing for the flexible integration of variables such as year, month, and specific four-hour intervals. Additionally, the Bayesian framework provides posterior distributions, offering deeper insights into parameter uncertainty and variability, which are critical for policy decisions. Therefore, Bayesian Poisson regression is more suitable for this study than the ARMA model.

@@

```{r, fig.pos="H"}
#| label: fig-PPplot
#| fig-cap: Posterior Predictive Checks
#| echo: false
#| message: false
#| warning: false
# 5. Posterior Predictive Checks
pp_check(poisson_glm, plotfun = "dens_overlay", nreps = 100) +
  labs(
    x = "Usage",  
    y = "Estimated density"         
  ) +
   yaxis_text()
```
The credible intervals shown in @tbl-crediinte highlight several strengths of the model. Key variables such as hour, month, and year have narrow credible intervals that do not include zero, indicating their significant and consistent influence on the response variable. This demonstrates the model’s ability to effectively capture critical trends and relationships in the data, particularly those related to time and seasonal effects. Additionally, the relatively small width of these intervals reflects the stability and reliability of parameter estimates, suggesting that the model is well-calibrated and robust in its predictions. The intercept, while having a broader credible interval, provides a reasonable baseline estimate, reflecting the overall level of the response variable.

@@

```{r, fig.pos="H", fig.width=8, fig.height=6}
#| label: fig-result
#| fig-cap: "The actual count from 4:00 AM to 8:00 AM on September 26, 2024 vs the predicted count for the same time interval in 2025."
#| echo: false
#| message: false
#| warning: false
ut_stations <- c("Madison Ave / Bloor St W", "Bloor St W / Huron St", "St. George St / Bloor St W", 
                 "Sussex Ave / St George St", "Spadina Ave / Sussex Ave", "Spadina Ave / Harbord St - SMART",
                 "St. George St / Hoskin Ave", "Spadina Ave / Willcocks St", "St. George St / Willcocks St",
                 "Willcocks St / St. George St", "Queen's Park / Bloor St W", "Queen's Park Cres W / Hoskin Ave",
                 "Wellesley St W / Queen's Park Cres", "Queen's Park Cres E / Grosvenor St - SMART", 
                 "Bay St / Bloor St W (East Side)", "Bay St / Bloor St W (West Side)", "Bay St / Charles St W - SMART",
                 "St. Joseph St / Bay St - SMART", "Bay St / St. Joseph St", "Bay St / Wellesley St W", 
                 "Ursula Franklin St / Huron St - SMART", "Ursula Franklin St / St. George St - SMART", "Galbraith Rd / King's College Rd",
                 "College St / Huron St", "College St / Henry St ", "Queens Park Cres / College St ", "University Ave / College St (East)")

@@

for (station in ut_stations) {
  station_data <- subset(start, station_name == station)
  path <- paste(paste0("models/", gsub("[ /]", "_", station), ".rds"))
  md <- readRDS(here::here(path))
  target <- data_frame(
    hour = 4,
    day = 26,
    month = 9,
    year = 2026
  )
  target$predicted <- ceiling(exp(predict(md, newdata = target)))
  target$station_name <- station
  target <- target %>% select(-hour,-day,-month,-year)
  result[result$station_name == station, "prediction"] <- target$predicted
}

data2024 <- start
for (station in ut_stations) {
  date <- "2024-09-26 04:00:00"
  exists <- nrow(data2024[data2024$station_name == station & data2024$time == date, ]) > 0

@@

names(result)[names(result) == "time"] <- "target_time"

data_fi <- merge(data2024, result, by = "station_name")

data_fi$growth <- data_fi$prediction - data_fi$count

@@

This study identifies significant patterns in the use of bike-sharing systems on the University of Toronto’s St. George campus, including temporal, seasonal, and annual trends. The findings reveal that 79.8% of bike usage occurs between 8 AM and 8 PM, while demand between midnight and 8 AM is minimal. This highlights the strong alignment between bike usage and users’ daily routines. To improve commuting efficiency, it is recommended that the university collaborates with the bike-sharing provider to increase bike availability during peak hours (e.g., before and after class times) and optimize station layouts, particularly in high-demand areas such as dormitories, academic buildings, and transit hubs. During low-demand hours (e.g., early morning), maintenance and redistribution efforts can be scheduled to ensure the availability of bikes during peak periods.

Seasonal data also show that bike usage significantly decreases during winter, approaching zero, while increasing steadily during spring and summer, peaking in August and September during the start of the academic year. This suggests the need for proactive and dynamic adjustments in bike supply. During the start of the academic year, expanding station capacity and offering user-friendly navigation services can help new and returning students quickly familiarize themselves with the system. In winter, efforts should focus on improving station infrastructure (e.g., anti-freeze facilities) and maintaining safety under cold conditions. Additionally, by leveraging predictive analytics to forecast station-specific demand fluctuations, resources can be optimized in advance, minimizing inconveniences caused by bike shortages or overcrowded parking during peak times.

## This Study Provides a Research Methodology for Shared Bikes in Semi-Closed Areas

@@

In these environments, bike-sharing systems can be optimized to align with peak demand periods, such as shift changes in factories, scheduled breaks in office parks, or recreational times in residential areas. By leveraging predictive analytics and data-driven insights, administrators can strategically allocate bikes and docking stations to match demand patterns, ensuring users have access when and where they need it most.

@@

The study only considers temporal variables (e.g., year, month, day, and hour) and excludes other critical external factors, significantly limiting the accuracy of its predictions. Weather conditions, for instance, have a substantial impact on bike usage; rainy, snowy, or extremely hot or cold days see a marked decline in usage frequency. Moreover, special events on campus and holidays directly affect bike demand. Neglecting these external factors prevents the model from accurately capturing fluctuations in actual demand.

The handling of temporal variables in the model is relatively simplistic, overlooking potential complex interactions among these variables, which restricts the predictive performance. Poisson regression struggles to adequately address the high concentration of demand during peak periods, potentially underestimating peak usage. Consequently, more sophisticated and flexible models, such as random forests or deep learning approaches, may be better suited to capturing these intricate relationships and improving prediction accuracy.

@@

## Suggestions for Future Research
While this study successfully predicts bike-sharing usage at the University of Toronto's St. George campus, several areas warrant further exploration to enhance the reliability and applicability of the results.

First, expanding the dataset to include additional environmental variables could provide deeper insights. Incorporating factors such as weather conditions, traffic congestion data, and socioeconomic variables could improve the model's accuracy, especially for predicting daily or hourly usage variations. Future studies could also include other campus transportation modes to enable a more comprehensive analysis of mobility patterns within the campus.

Additionally, this research is limited to employing Bayesian Poisson regression. Exploring more advanced machine learning methods, such as random forests or neural networks, could improve prediction accuracy by capturing the nonlinear relationships among factors influencing bike-sharing usage.

@@

# Data Collection Methodology
## Overview
Bike Share Toronto is the largest bike-sharing service provider in the Greater Toronto Area, committed to promoting gas-free transportation since its establishment. Starting with 79 stations in 2014, the system has expanded rapidly over the past decade. As of September 2024, the number of stations has grown to 855, capturing the entire bike-sharing market in Toronto [@liu2023bike].

Bike Share Toronto uses its advanced bike-sharing system to automatically collect and manage ride data. Each bike and docking station is equipped with sensors and communication devices to record trip data in detail. This data is transmitted in real time to a central database via a wireless network, used for system monitoring, operational maintenance, and user behavior analysis. To support academic research and urban planning, Bike Share Toronto also anonymizes the data and makes it publicly available. This dataset includes all ride records since the system's inception and is therefore considered a population rather than a sample.

@@

**Completeness**  
The dataset covers all ride records since 2014, with data variables expanding as the system and technology improved. This historical and comprehensive nature allows the data to reflect long-term trends in Toronto's bike-sharing system. However, the limited variables in early data (e.g., only hourly usage for 2014-2015) require researchers to be mindful of inconsistencies across years, and early data may need supplementation or transformation to meet research needs.

@@

**Relevance**  
The detailed variables in the dataset (such as station ID, ride duration, and user type) provide opportunities for multidimensional analysis, enabling researchers to explore geographic distribution, temporal patterns, and user behavior. Since the dataset represents a population rather than a sample, researchers have greater flexibility in their study focus, ranging from micro-level user behavior to macro-level trend analysis. However, this comprehensiveness also increases the difficulty of data cleaning, such as dealing with redundant data, outliers, and the computational burden of large-scale data processing. Moreover, anonymization protects user privacy, it may limit certain studies focusing on micro-level user behavior.

@@

Diversity in data collection is crucial for enhancing research flexibility and ensuring reliable results. A well-rounded dataset for bike-sharing systems should consider several key aspects:

@@

Ensuring **compatibility of data formats and platforms** is critical for effective use of diverse datasets. Collected data should support structured formats (e.g., rental records), unstructured formats (e.g., user feedback), and geospatial information (e.g., GIS formats). Standardized formats like CSV, JSON, or Parquet along with metadata should be used for easy integration across platforms such as GIS systems, machine learning tools, and statistical analysis tools.

@@

Implementing **redundant data collection mechanisms** is crucial for preventing data loss due to sensor malfunctions. Installing multiple sensors on bikes and stations serves as a backup system. For instance, sensors on bikes can complement the data from docking stations. A centralized monitoring system can cross-check data from multiple sources to detect inconsistencies or missing information, thereby improving data integrity and reliability.

To enhance data comprehensiveness, integrating **authoritative environmental and traffic data APIs** can be highly beneficial. For example, APIs from OpenWeatherMap or government meteorological services can provide real-time information on weather conditions, air quality, wind speed, and precipitation. Traffic data from relevant authorities can provide insights into road congestion levels. Combining these variables with bike-sharing records can reveal how external conditions affect user behavior, ultimately guiding bike allocation and station optimization.

@@

Ensuring **data transparency** is crucial. All corrected records should be labeled with information about the sources, methods, and reasons for modification. This transparency helps maintain data integrity and fosters trust among researchers and system operators.

@@

```{r}
#| warning: false
#| message: false
#| echo: false
#| label: tbl-p2
#| tbl-cap: "Simulated Ideal Dataset Part 2"
bike_share_part2 <- data.frame(
  End_Station_ID = sample(100:200, 5, replace = TRUE),
  End_Station_Name = paste("Station", sample(100:200, 5, replace = TRUE)),
  Latitude = round(runif(5, 43.6, 43.7), 6),
  Longitude = round(runif(5, -79.4, -79.3), 6),
  Trip_ID = paste("TRIP", sample(1000:9999, 5, replace = TRUE), sep = "-")
)

@@

```{r}
#| warning: false
#| message: false
#| echo: false
#| label: tbl-p3
#| tbl-cap: "Simulated Ideal Dataset Part 3"
bike_share_part3 <- data.frame(
  Trip_Duration_s = sample(300:3600, 5, replace = TRUE),
  Trip_Distance_m = sample(500:10000, 5, replace = TRUE),
  Bike_ID = paste("BIKE", sample(1000:9999, 5, replace = TRUE), sep = "-"),
  User_Type = sample(c("Member", "Casual"), 5, replace = TRUE)
)

@@

```{r}
#| warning: false
#| message: false
#| echo: false
#| label: tbl-p4
#| tbl-cap: "Simulated Ideal Dataset Part 4"
bike_share_part4 <- data.frame(
  Weather_Conditions = sample(c("Sunny", "Cloudy", "Rainy"), 5, replace = TRUE),
  Air_Quality_Index = sample(10:100, 5, replace = TRUE),
  Traffic_Conditions = sample(c("Low", "Moderate", "High"), 5, replace = TRUE),
  Available_Bikes_at_Start_Station = sample(0:20, 5, replace = TRUE)
)

@@

```{r}
#| warning: false
#| message: false
#| echo: false
#| label: tbl-p5
#| tbl-cap: "Simulated Ideal Dataset Part 5"
bike_share_part5 <- data.frame(
  Available_Docks_at_End_Station = sample(0:20, 5, replace = TRUE),
  Device_Status = sample(c("Operational", "Faulty"), 5, replace = TRUE),
  Maintenance_Records = sample(c("None", "Minor Repair", "Major Repair"), 5, replace = TRUE)
)

@@

```{r}
#| warning: false
#| message: false
#| echo: false
#| label: tbl-p6
#| tbl-cap: "Simulated Ideal Dataset Part 6"
bike_share_part6 <- data.frame(
  User_Gender = sample(c("Male", "Female", "Other"), 5, replace = TRUE),
  Anomalies = sample(c("None", "Sensor Error", "Docking Failure"), 5, replace = TRUE),
  Missing_Data_Flag = sample(c(TRUE, FALSE), 5, replace = TRUE),
  Data_Source = sample(c("Station Sensor", "Mobile App", "System Log"), 5, replace = TRUE)
)

>>>>

## Overview
Bike-sharing services have experienced growth in the Toronto region, with ridership rising dramatically from approximately 665,000 trips in 2015 to over 4.5 million in 2022, according to a study by the University of Toronto's School of Cities[@liu2023bike]. This increase highlights the need for analyses to optimize bike-sharing systems and guide policy decisions that can further enhance urban mobility[@xiamen]. While city-wide studies provide suggestions, research from Hangzhou, China, has demonstrated that bike-sharing usage patterns can vary significantly between urban areas and university campuses[@hangzhou]. These differences suggest that general urban studies conducted in the Toronto region may not fully capture the unique dynamics of campus-based bike-sharing systems. Therefore, to address this unexplored areas in the literature, this study focuses on the University of Toronto St. George Campus as a case study, analyzing bike-sharing usage within the campus in Toronto to provide a detailed understanding of its role in campus mobility and inform policies tailored to similar settings.

@@

This data is particularly ideal for the study also due to its high reliability, which stems from the commercial nature of bike-sharing operations. The bike-sharing system ensures accurate data collection and management by automatically recording ride information. Each bike and docking station is equipped with sensors and communication devices, enabling the capture of detailed trip data.

@@

#Display the dataset.
start |>
  tail() |>
  kable(booktabs = TRUE, align = "c")
```

@@

@tbl-varibale is the summary of varibales. It is evident that the dataset includes 27 unique station names and spans nearly eight years (from January 1, 2017, to September 30, 2024), providing opportunities to analyze long-term trends and seasonal variations. In the count column, the minimum and median values are both 1, the mean is 2, the first quartile is 1, the third quartile is 3, and the maximum reaches 31, showing a significant right-skewed distribution. This indicates that most of the records have low count values, but there are a few extreme high values.
```{r}
#| warning: false
#| message: false
#| echo: false
#| label: tbl-varibale
#| tbl-cap: "Variables' summary"

#Calculate the summary of each variable.
count_min <- min(start$count)
count_1st_qu <- quantile(start$count, 0.25)
count_median <- median(start$count)
count_mean <- mean(start$count)
count_3rd_qu <- quantile(start$count, 0.75)
count_max <- max(start$count)

@@

# Generate a copy of the data.
start_asdate <- start 
start_asdate$time <- as.Date(start_asdate$time)

# Calculate the total count for each day.
daily_counts <- start_asdate %>%
  group_by(time) %>%
  summarise(total_count = sum(count))

#Calculate the slope.
lm_model <- lm(total_count ~ time, data = daily_counts)
slope <- coef(lm_model)["time"]

@@

# Generate a copy of the data.
start_astime <- start 
start_astime$time <- as.POSIXct(start_astime$time)

# Calculate the total count for each time period.
start_astime <- start_astime %>%
  mutate(time_interval = case_when(
    hour(time) >= 0 & hour(time) < 4 ~ "00:00 - 04:00",
    hour(time) >= 4 & hour(time) < 8 ~ "04:00 - 08:00",
    hour(time) >= 8 & hour(time) < 12 ~ "08:00 - 12:00",
    hour(time) >= 12 & hour(time) < 16 ~ "12:00 - 16:00",
    hour(time) >= 16 & hour(time) < 20 ~ "16:00 - 20:00",
    hour(time) >= 20 & hour(time) < 24 ~ "20:00 - 00:00"
  ))

#Calculate the proportion.
total_counts <- start_astime %>%
  group_by(time_interval) %>%
  summarise(count = sum(count, na.rm = TRUE)) %>%
  mutate(percentage = count / sum(count) * 100)

@@

# Calculate the daily average.
average_counts_start <- start_1day1 %>%
  group_by(station_name) %>%
  summarize(average_count = mean(total_count, na.rm = TRUE))
average_counts_stop <- stop_1day1 %>%
  group_by(station_name) %>%
  summarize(average_count = mean(total_count, na.rm = TRUE))

@@

# Handle non-positive integer mean.
merged_data <- merged_data %>%
  mutate(count_difference = ifelse((average_count.y - average_count.x) >= 0, 
                                   ceiling(average_count.y - average_count.x), 
                                   floor(average_count.y - average_count.x)))

@@

This study considered using the ARIMA (autoregressive integrated moving average) model as an alternative but ultimately decided against it due to its limitations in capturing the specific characteristics of bike-sharing usage data. Although the ARIMA model has advantages in time series analysis, allowing for more accurate short-term predictions with parameters that have clearer interpretability. Bike-sharing usage exhibits significant seasonal patterns, time-of-day effects, and count-based characteristics, while ARIMA models are primarily suited for stationary time series with linear autocorrelations. ARIMA struggles to effectively address non-stationary trends and overdispersion, such as the marked decrease in usage during winter. Although extensions like SARIMA (Seasonal ARIMA) can account for seasonality, they still cannot directly handle the count-based nature of the data. In contrast, Bayesian Poisson regression is specifically designed for count data, allowing for the flexible integration of variables such as year, month, and specific four-hour intervals. Additionally, the Bayesian framework provides posterior distributions, offering deeper 
ideas into parameter uncertainty and variability, which are 
important for policy decisions. Therefore, Bayesian Poisson regression is more suitable for this study than the ARMA model.

@@

```{r, fig.pos="H"}
#| label: fig-PPplot
#| fig-cap: Posterior Predictive Checks
#| echo: false
#| message: false
#| warning: false
# Posterior Predictive Checks
pp_check(poisson_glm, plotfun = "dens_overlay", nreps = 100) +
  labs(
    x = "Usage",  
    y = "Estimated density"         
  ) +
   yaxis_text()
```
The credible intervals shown in @tbl-crediinte highlight several strengths of the model. Key variables such as hour, month, and year have narrow credible intervals that do not include zero, indicating their significant and consistent influence on the response variable. This demonstrates the model’s ability to effectively capture important trends and relationships in the data, particularly those related to time and seasonal effects. Additionally, the relatively small width of these intervals reflects the stability and reliability of parameter estimates, suggesting that the model is well-calibrated and robust in its predictions. The intercept, while having a broader credible interval, provides a reasonable baseline estimate, reflecting the overall level of the response variable.

@@

```{r, fig.pos="H", fig.width=8, fig.height=6}
#| label: fig-result
#| fig-cap: "The actual count from 4:00 AM to 8:00 AM on September 26, 2024 vs the predicted count for the same time interval in 2025."
#| echo: false
#| message: false
#| warning: false

# Target stations
ut_stations <- c("Madison Ave / Bloor St W", "Bloor St W / Huron St", "St. George St / Bloor St W", 
                 "Sussex Ave / St George St", "Spadina Ave / Sussex Ave", "Spadina Ave / Harbord St - SMART",
                 "St. George St / Hoskin Ave", "Spadina Ave / Willcocks St", "St. George St / Willcocks St",
                 "Willcocks St / St. George St", "Queen's Park / Bloor St W", "Queen's Park Cres W / Hoskin Ave",
                 "Wellesley St W / Queen's Park Cres", "Queen's Park Cres E / Grosvenor St - SMART", 
                 "Bay St / Bloor St W (East Side)", "Bay St / Bloor St W (West Side)", "Bay St / Charles St W - SMART",
                 "St. Joseph St / Bay St - SMART", "Bay St / St. Joseph St", "Bay St / Wellesley St W", 
                 "Ursula Franklin St / Huron St - SMART", "Ursula Franklin St / St. George St - SMART", "Galbraith Rd / King's College Rd",
                 "College St / Huron St", "College St / Henry St ", "Queens Park Cres / College St ", "University Ave / College St (East)")

@@

# Iterate through all models.
for (station in ut_stations) {
  station_data <- subset(start, station_name == station)
  path <- paste(paste0("models/", gsub("[ /]", "_", station), ".rds"))
  md <- readRDS(here::here(path))
  target <- data_frame(
    hour = 4,
    day = 26,
    month = 9,
    year = 2026
  )
  target$predicted <- ceiling(exp(predict(md, newdata = target)))
  target$station_name <- station
  target <- target %>% select(-hour,-day,-month,-year)
  result[result$station_name == station, "prediction"] <- target$predicted
}

data2024 <- start

# Find predictions.
for (station in ut_stations) {
  date <- "2024-09-26 04:00:00"
  exists <- nrow(data2024[data2024$station_name == station & data2024$time == date, ]) > 0

@@

# Rename
names(result)[names(result) == "time"] <- "target_time"

# New dataset for ploting.
data_fi <- merge(data2024, result, by = "station_name")

# Calculate predicted growth
data_fi$growth <- data_fi$prediction - data_fi$count

@@

This study identifies significant patterns in the use of bike-sharing systems on the University of Toronto’s St. George campus, including temporal, seasonal, and annual trends. The findings shows that 79.8% of bike usage occurs between 8 AM and 8 PM, while demand between midnight and 8 AM is minimal. This highlights the strong alignment between bike usage and users’ daily routines. To improve commuting efficiency, it is recommended that the university collaborates with the bike-sharing provider to increase bike availability during peak hours (e.g., before and after class times) and optimize station layouts, particularly in high-demand areas such as dormitories, academic buildings, and transit hubs. During low-demand hours (e.g., early morning), maintenance and redistribution efforts can be scheduled to ensure the availability of bikes during peak periods.

Seasonal data also show that bike usage significantly decreases during winter, approaching zero, while increasing steadily during spring and summer, peaking in August and September during the start of the academic year. This suggests the need for proactive and dynamic adjustments in bike supply. During the start of the academic year, expanding station capacity and offering user-friendly navigation services can help new and returning students quickly familiarize themselves with the system. In winter, efforts should focus on improving station infrastructure (e.g., anti-freeze facilities) and maintaining safety under cold conditions. Additionally, by utilizing predictive analytics to forecast station-specific demand fluctuations, resources can be optimized in advance, minimizing inconveniences caused by bike shortages or overcrowded parking during peak times.

## A New Research Methodology for Shared Bikes in Semi-Closed Areas

@@

In these environments, bike-sharing systems can be optimized to align with peak demand periods, such as shift changes in factories, scheduled breaks in office parks, or recreational times in residential areas. By leveraging predictive analytics administrators can strategically allocate bikes and docking stations to match demand patterns, ensuring users have access when and where they need it most.

@@

The study only considers temporal variables (e.g., year, month, day, and hour) and excludes other important external factors, significantly limiting the accuracy of its predictions. Weather conditions, for instance, have a substantial impact on bike usage; rainy, snowy, or extremely hot or cold days see a marked decline in usage frequency. Moreover, special events on campus and holidays directly affect bike demand. Neglecting these external factors prevents the model from accurately capturing fluctuations in actual demand.

The handling of temporal variables in the model is relatively simplistic, overlooking potential complex interactions among these variables, which restricts the predictive performance. Poisson regression struggles to adequately address the high concentration of demand during peak periods, potentially underestimating peak usage. Consequently, more sophisticated and flexible models, such as random forests or deep learning approaches, may be better suited to capturing these complicated relationships and improving prediction accuracy.

@@

## Suggestions for Future Research
While this study successfully predicts bike-sharing usage at the University of Toronto's St. George campus, several areas warrant further research to enhance the reliability and applicability of the results.

First, expanding the dataset to include additional environmental variables could provide deeper understandings. Incorporating factors such as weather conditions, traffic congestion data, and socioeconomic variables could improve the model's accuracy, especially for predicting daily or hourly usage variations. Future studies could also include other campus transportation modes to enable a more detailed analysis of mobility patterns within the campus.

Additionally, this research is limited to employing Bayesian Poisson regression. Exploring more machine learning methods, such as random forests or neural networks, could improve prediction accuracy by capturing the nonlinear relationships among factors influencing bike-sharing usage.

@@

# Data Collection Methodology
## Overview
Bike Share Toronto is the largest bike-sharing service provider in the Greater Toronto Area, committed to promoting gas-free transportation since its establishment. Starting with 79 stations in 2014, the system has expanded over the past decade. As of September 2024, the number of stations has grown to 855, capturing the entire bike-sharing market in Toronto [@liu2023bike].

Bike Share Toronto uses its bike-sharing system to automatically collect and manage ride data. Each bike and docking station is equipped with sensors and communication devices to record trip data in detail. This data is transmitted in real time to a central database via a wireless network, used for system monitoring, operational maintenance, and user behavior analysis. To support academic research and urban planning, Bike Share Toronto also anonymizes the data and makes it publicly available. This dataset includes all ride records since the system's inception and is therefore considered a population rather than a sample.

@@

**Completeness**  
The dataset covers all ride records since 2014, with data variables expanding as the system and technology improved. This historical and detailed nature allows the data to reflect long-term trends in Toronto's bike-sharing system. However, the limited variables in early data (e.g., only hourly usage for 2014-2015) require researchers to be mindful of inconsistencies across years, and early data may need supplementation or transformation to meet research needs.

@@

**Relevance**  
The detailed variables in the dataset (such as station ID, ride duration, and user type) provide opportunities for multidimensional analysis, enabling researchers to explore geographic distribution, temporal patterns, and user behavior. Since the dataset represents a population rather than a sample, researchers have greater flexibility in their study focus, ranging from micro-level user behavior to macro-level trend analysis. However, this exhaustiveness also increases the difficulty of data cleaning, such as dealing with redundant data, outliers, and the computational burden of large-scale data processing. Moreover, anonymization protects user privacy, it may limit certain studies focusing on micro-level user behavior.

@@

Diversity in data collection is important for enhancing research flexibility and ensuring reliable results. A well-rounded dataset for bike-sharing systems should consider several key aspects:

@@

Ensuring **compatibility of data formats and platforms** is important for effective use of diverse datasets. Collected data should support structured formats (e.g., rental records), unstructured formats (e.g., user feedback), and geospatial information (e.g., GIS formats). Standardized formats like CSV, JSON, or Parquet along with metadata should be used for easy integration across platforms such as GIS systems, machine learning tools, and statistical analysis tools.

@@

Implementing **redundant data collection mechanisms** is important for preventing data loss due to sensor malfunctions. Installing multiple sensors on bikes and stations serves as a backup system. For instance, sensors on bikes can complement the data from docking stations. A centralized monitoring system can cross-check data from multiple sources to detect inconsistencies or missing information, thereby improving data integrity and reliability.

To enhance data exhaustiveness, integrating **authoritative environmental and traffic data APIs** can be highly beneficial. For example, APIs from OpenWeatherMap or government meteorological services can provide real-time information on weather conditions, air quality, wind speed, and precipitation. Traffic data from relevant authorities can provide understandings into road congestion levels. Combining these variables with bike-sharing records can understand how external conditions affect user behavior, ultimately guiding bike allocation and station optimization.

@@

Ensuring **data transparency** is important. All corrected records should be labeled with information about the sources, methods, and reasons for modification. This transparency helps maintain data integrity and fosters trust among researchers and system operators.

@@

```{r}
#| warning: false
#| message: false
#| echo: false
#| label: tbl-p2
#| tbl-cap: "Simulated Ideal Dataset Part 2"

#Simulated Ideal Dataset Part 2.
bike_share_part2 <- data.frame(
  End_Station_ID = sample(100:200, 5, replace = TRUE),
  End_Station_Name = paste("Station", sample(100:200, 5, replace = TRUE)),
  Latitude = round(runif(5, 43.6, 43.7), 6),
  Longitude = round(runif(5, -79.4, -79.3), 6),
  Trip_ID = paste("TRIP", sample(1000:9999, 5, replace = TRUE), sep = "-")
)

@@

```{r}
#| warning: false
#| message: false
#| echo: false
#| label: tbl-p3
#| tbl-cap: "Simulated Ideal Dataset Part 3"
#Simulated Ideal Dataset Part 3.
bike_share_part3 <- data.frame(
  Trip_Duration_s = sample(300:3600, 5, replace = TRUE),
  Trip_Distance_m = sample(500:10000, 5, replace = TRUE),
  Bike_ID = paste("BIKE", sample(1000:9999, 5, replace = TRUE), sep = "-"),
  User_Type = sample(c("Member", "Casual"), 5, replace = TRUE)
)

@@

```{r}
#| warning: false
#| message: false
#| echo: false
#| label: tbl-p4
#| tbl-cap: "Simulated Ideal Dataset Part 4"
#Simulated Ideal Dataset Part 4.
bike_share_part4 <- data.frame(
  Weather_Conditions = sample(c("Sunny", "Cloudy", "Rainy"), 5, replace = TRUE),
  Air_Quality_Index = sample(10:100, 5, replace = TRUE),
  Traffic_Conditions = sample(c("Low", "Moderate", "High"), 5, replace = TRUE),
  Available_Bikes_at_Start_Station = sample(0:20, 5, replace = TRUE)
)

@@

```{r}
#| warning: false
#| message: false
#| echo: false
#| label: tbl-p5
#| tbl-cap: "Simulated Ideal Dataset Part 5"
#Simulated Ideal Dataset Part 5.
bike_share_part5 <- data.frame(
  Available_Docks_at_End_Station = sample(0:20, 5, replace = TRUE),
  Device_Status = sample(c("Operational", "Faulty"), 5, replace = TRUE),
  Maintenance_Records = sample(c("None", "Minor Repair", "Major Repair"), 5, replace = TRUE)
)

@@

```{r}
#| warning: false
#| message: false
#| echo: false
#| label: tbl-p6
#| tbl-cap: "Simulated Ideal Dataset Part 6"
#Simulated Ideal Dataset Part 6.
bike_share_part6 <- data.frame(
  User_Gender = sample(c("Male", "Female", "Other"), 5, replace = TRUE),
  Anomalies = sample(c("None", "Sensor Error", "Docking Failure"), 5, replace = TRUE),
  Missing_Data_Flag = sample(c(TRUE, FALSE), 5, replace = TRUE),
  Data_Source = sample(c("Station Sensor", "Mobile App", "System Log"), 5, replace = TRUE)
)
